The impact of environmental shocks on children's schooling in Bolivia

### 2020-05-07 paper_draft.Rmd
added interaction terms

### Variable descriptions

- **hh_id:** The household ID, which is the same for each member of the household. It is made up of letters and numbers. 
- **nro1:** The number of the person in each household.
- **assist:** Binary variable, whether or not the individual assists school 
- **urb_rur:** Binary variable, whether the household is in an urban or rural area
- **reason:** the reason why the individual is not assisting school if they are not 
- **members:** number of household members excluding servants 
- **sex:** the sex of the individual, Male = 1, Female =2 
- **age:** How old the individual is, in years 
- **relation:** The relationship of the individual to the head of household 
- **literacy:** Binary variable, whether a person can read and write 
- **register:** Binary variable, whether an individual registered for school or not 
- **grade:** For what grade is the person registered 
- **repeated:** whether the person has repeated the grade they are currently registered in 
- **dept:** the department in which the person lives 
- **income:** the per capita monthly income for each household, in bolivianos

@@

hh_2005 <- data_2005 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2005,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
         reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         assist = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 17 ~ 1,
                             school_level != 17 ~ 0),
         secondary = case_when(school_level == 18 ~ 1,
                               school_level != 18 ~ 0))

@@

hh_2006 <- data_2006 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2006,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         assist = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 17 ~ 1,
                             school_level != 17 ~ 0),
         secondary = case_when(school_level == 18 ~ 1,
                               school_level != 18 ~ 0))
```

@@

hh_2007 <- data_2007 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2007,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         assist = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 17 ~ 1,
                             school_level != 17 ~ 0),
         secondary = case_when(school_level == 18 ~ 1,
                               school_level != 18 ~ 0))
```

@@

hh_2011 <- data_2011 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2011,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                               literacy1 == 2 ~ 0), 
          female = case_when(sex == 2 ~ 1,
                             sex == 1 ~ 0),
           rural = case_when(urb_rur == 1 ~ 0,
                             urb_rur == 2 ~ 1), 
           assist = case_when(assist1 == 1 ~ 1,
                              assist1 == 2 ~ 0),
           register = case_when(register1 == 1 ~ 1,
                                register1 == 2 ~ 0),
         primary = case_when(school_level == 16 ~ 1,
                             school_level != 16 ~ 0),
         secondary = case_when(school_level == 17 ~ 1,
                               school_level != 17 ~ 0))
```

@@

hh_2012 <- data_2012 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2012,
         dept = case_when(depto== "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
         literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         assist = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 19 ~ 1,
                             school_level != 19 ~ 0),
         secondary = case_when(school_level == 20 ~ 1,
                               school_level != 20 ~ 0))

@@

# Data Processing and Combination

To create a complete dataset, I merged the household data with the floods data and the droughts data. Before merging the data, I cut down the datasets to include only the relevant variables needed for analysis in each file. I renamed several variables in the household survey data set to make them more understandable. I also created the year variable for the household data to ensure that it matched up correctly with the weather data. For the flood and drought data, I grouped the occurrences by department instead of by municipality so they would match the household data, and I made the department character numeric for the merge.

@@

total_data %>%
  group_by(work_reason, money_reason) %>%
  summarise(n = n())

@@

children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "female", "child", "grandchild", "literacy", "dept", "rural", "assist", "reason", "register", "primary", "secondary", "floods", "droughts", "year", "job", "tasks", "money_reason", "work_reason"))

@@

#total_2 <- total[!(total$reason == "vacation" & total$assist == 0]

@@

total_2 <- subset(total, !(reason == "vacation" & assist == 0)) #taking out the observations with vacation kids

@@

```
```{r summary statistic tables with total_subset data, echo = F, message = F}
#creating a table with the main explanatory variables

#creating a table dividing the characteristics by urban/rural area
```
#Empirical Model
see the link for math equation help.

@@

reg_data <- total_subset %>%
  mutate(year_factor = as.factor(year),
         dept_factor = as.factor(dept),
         lny = log(income + 1),
         valley = case_when(dept == "Chuquisaca", "Cochabamba", "Tarija" ~ 1,
                            dept != "Chuquisaca", "Cochabamba", "Tarija" ~ 0))

#adding the region variables for the region fixed-effects option

reg_data %>%
  group_by(dept_factor) %>%
  summarise(n = n())

@@

@@

#model_simple <- lm(assist ~ floods + droughts + year, data = reg_data)
#summary(model_simple)
model_assist <- lm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
summary(model_assist)
model_register <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
summary(model_register)
model_reason_t <- (lm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data))
summary(model_reason_t)

@@

```{r generate regression table, echo = F, results = "asis"}
stargazer::stargazer(list(model_assist, model_register), 
               title = "Full sample regression results", 
               label = "fullResults",
               header = F,
               keep = c("floods", "droughts"))
```
```{r probit regression total sample, echo = F, message = F}

probitMod_assist <- glm(assist ~ floods + droughts + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))

@@

```{r generate probit regression table, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist, probitMod_register, probitMod_work), 
               title = "Full sample probit regression results", 
               label = "fullResults",
               header = F,
               keep = c("floods", "droughts"))

@@

rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "primary"))

model_assist_r <- (lm(assist ~ floods + droughts + age + female + income + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
summary(model_assist_r)
model_register_r <- lm(register ~ floods + droughts + age + female + income + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
summary(model_register_r)
model_reason <- (lm(work_reason ~ floods + droughts + age + female + income + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
summary(model_reason)

@@

```{r generate linear regression table rural sample, echo = F, results = "asis"}
stargazer::stargazer(list(model_assist, model_register), 
               title = "Rural sample linear regression results", 
               label = "fullResults",
               header = F,
               keep = c("floods", "droughts"))
```

@@

probitMod_assist_r <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))

@@

urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor"))

model_assist_u <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
summary(model_assist_u)
model_register_u <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
summary(model_register_u)
```

@@

probitMod_assist_u <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))

>>>>

@@

hh_2005 <- data_2005 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2005,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
         reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         attend = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 17 ~ 1,
                             school_level != 17 ~ 0),
         secondary = case_when(school_level == 18 ~ 1,
                               school_level != 18 ~ 0))

@@

hh_2006 <- data_2006 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2006,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         attend = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 17 ~ 1,
                             school_level != 17 ~ 0),
         secondary = case_when(school_level == 18 ~ 1,
                               school_level != 18 ~ 0))
```

@@

hh_2007 <- data_2007 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2007,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         attend = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 17 ~ 1,
                             school_level != 17 ~ 0),
         secondary = case_when(school_level == 18 ~ 1,
                               school_level != 18 ~ 0))
```

@@

hh_2011 <- data_2011 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2011,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
          literacy = case_when(literacy1 == 1 ~ 1,
                               literacy1 == 2 ~ 0), 
          female = case_when(sex == 2 ~ 1,
                             sex == 1 ~ 0),
           rural = case_when(urb_rur == 1 ~ 0,
                             urb_rur == 2 ~ 1), 
           attend = case_when(assist1 == 1 ~ 1,
                              assist1 == 2 ~ 0),
           register = case_when(register1 == 1 ~ 1,
                                register1 == 2 ~ 0),
         primary = case_when(school_level == 16 ~ 1,
                             school_level != 16 ~ 0),
         secondary = case_when(school_level == 17 ~ 1,
                               school_level != 17 ~ 0))
```

@@

hh_2012 <- data_2012 %>% 
  select(hh_id, assist1, reason, sex, age,
         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
  mutate(year = 2012,
         dept = case_when(depto== "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"),
           reason = case_when(reason == 1 ~ 'vacation',
                            reason == 2 ~ 'money', 
                            reason == 3 ~ 'work', 
                            reason == 4 ~ 'sickness/accident/disability', 
                            reason == 5 ~ 'distance',
                            reason == 6 ~ 'finished', 
                            reason == 7 ~ 'age',
                            reason == 8 ~ 'not interested', 
                            reason == 9 ~ 'housework/childcare',
                            reason == 10 ~ 'other'),
         literacy = case_when(literacy1 == 1 ~ 1,
                              literacy1 == 2 ~ 0), 
         female = case_when(sex == 2 ~ 1,
                            sex == 1 ~ 0),
         rural = case_when(urb_rur == 1 ~ 0,
                           urb_rur == 2 ~ 1), 
         attend = case_when(assist1 == 1 ~ 1,
                            assist1 == 2 ~ 0),
         register = case_when(register1 == 1 ~ 1,
                              register1 == 2 ~ 0),
         primary = case_when(school_level == 19 ~ 1,
                             school_level != 19 ~ 0),
         secondary = case_when(school_level == 20 ~ 1,
                               school_level != 20 ~ 0))

@@

@@

#total_data %>%
  #group_by(work_reason, money_reason) %>%
  #summarise(n = n())

@@

children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "female", "child", "grandchild", "literacy", "dept", "rural", "attend", "reason", "register", "primary", "secondary", "floods", "droughts", "year", "job", "tasks", "money_reason", "work_reason"))

@@

@@

total_2 <- subset(total, !(reason == "vacation" & attend == 0)) #taking out the observations with vacation kids

@@

@@

reg_data <- total_subset %>%
  mutate(year_factor = as.factor(year),
         dept_factor = as.factor(dept),
         lny = log(income + 1),
         valleys = case_when(dept %in% c("Chuquisaca", "Cochahamba", "Tarija") ~ 1, TRUE ~ 0),
         mountains = case_when(dept %in% c("La Paz", "Oruro", "Potosi") ~ 1, TRUE ~ 0),
         plains = case_when(dept %in% c("Beni", "Pando", "Santa Cruz") ~ 1, TRUE ~ 0),
         flood_secondary = floods * secondary,
         drought_secondary = droughts * secondary,
         flood_pov = floods * extr,
         drought_pov = droughts * extr)

#adding the region variables for the region fixed-effects option
#adding interaction terms
#adding lagged flood and drought variables

#creating rural sample
rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov"))

reg_data %>%
  group_by(dept_factor) %>%
  #summarise(n = n())

```
##Variable Descriptions

\begin{table}
\centering
\caption{Variable definitions}
\label{variableDefinitions}
\begin{tabular}{ll}
\hline
\hline
Variable name      & Description                                               \\
\hline
Attend          & Binary variable, 1 = attends school                          \\
Register        & Binary variable, 1 = registered for school                   \\
Work reason     & Binary variable, 1 = missing school for work                 \\
Lny             & Log of income per capita (in Bolivianos/month)               \\
Extr            & Binary variable, 1 = below extreme poverty line              \\
Female          & Binary variable, 1 = female                                  \\
Age             & Age in years                                                 \\
Child           & Binary variable, 1 = is child of head of household           \\
Grandchild      & Binary variable, 1 = is grandchild of head of household      \\
Num children    & Number of children in the household                          \\
Secondary       & Binary variable, 1 = in secondary school group               \\
hhhead educ     & Years of education of head of household                      \\
female hhhead   & Binary variable, 1 = female head of household                \\
Floods          & Number of flood occurrences per dept per year                \\
Droughts        & Number of drought occurrences per dept per year              \\
Dept            & Department                                                   \\
Valleys         & Binary variable, 1 = department is in valleys region         \\
Mountains       & Binary variable, 1 = department is in mountains region       \\
Plains          & Binary variable, 1 = department is in plains region          \\

\hline
\end{tabular}
\end{table}

```{r create summary statistic table, echo = F, results = 'asis'}
#creating a table with the main explanatory variables

summary_data <- reg_data %>%
  select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains)%>%
  summarise_all(list(mean = mean, sd = sd))

stargazer::stargazer(as.data.frame(summary_data),
                     label = "summaryStats",
                     title = "Summary statistics (full sample)",
                     header = F,
                     type = "latex",
                     digits = 0)

#creating a table dividing the characteristics by urban/rural area
summary_rural <- rural %>%
  select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains) %>%
  summarise_all(list(mean = mean, sd = sd))

stargazer::stargazer(as.data.frame(summary_data),
                     label = "summaryStatsR",
                     title = "Summary statistics (rural sample)",
                     header = F,
                     type = "latex",
                     digits = 0)

@@

#Empirical Model
see the link for math equation help.

@@

#model_simple <- lm(assist ~ floods + droughts + year, data = reg_data)
#summary(model_simple)
model_assist <- lm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
summary(model_assist)
model_register <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
summary(model_register)
model_reason_t <- (lm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data))
summary(model_reason_t)

@@

```{r generate regression table, echo = F, results = "asis"}
stargazer::stargazer(list(model_assist, model_register), 
               title = "Full sample regression results", 
               label = "fullResults",
               header = F,
               keep = c("floods", "droughts"))
```

```{r probit regression total sample, echo = F, message = F}

probitMod_assist <- glm(attend ~ floods + droughts + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))

@@

```{r generate probit regression table, echo = F, results = "asis"}
stargazer::stargazer(list(probitMod_assist, probitMod_register, probitMod_work), 
               title = "Full sample probit regression results", 
               label = "fullResults1",
               header = F,
               keep = c("floods", "droughts"))

@@

model_assist_r <- (lm(attend ~ floods + droughts + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
summary(model_assist_r)
model_register_r <- lm(register ~ floods + droughts + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
summary(model_register_r)
model_reason <- (lm(work_reason ~ floods + droughts + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
summary(model_reason)

@@

```{r generate linear regression table rural sample, echo = F, results = "asis"}
stargazer::stargazer(list(model_assist, model_register), 
               title = "Rural sample linear regression results", 
               label = "fullResults2",
               header = F,
               keep = c("floods", "droughts"))
```

@@

probitMod_assist_r <- glm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))

@@

urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "work_reason", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov"))

model_assist_u <- lm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
summary(model_assist_u)
model_register_u <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
summary(model_register_u)
```

@@

probitMod_assist_u <- glm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
